---
title: "generate_pvalue4CurrentManuscript"
date: "`r Sys.Date()`"
author: "Lu Yang & Jun Chen"
output:
 html_document:
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
      smooth_scroll: no
---



```{r, message=FALSE, message=F, warning=F, echo=FALSE, results='hide'}
grp.name= 'PR4.n'
adj.name= 'BRAF'
subsetBRAF= FALSE
BRAF.sub=''
```

```{r, message=FALSE, message=F, warning=F, echo=FALSE, results='hide'}
### Previously we use Lymphoctyes_KnitWithParams.Rmd to generate the html files
###########################################
pkg <- c('dplyr','tibble','tidyr','vegan','matrixStats', 'GUniFrac','ggpubr','RColorBrewer','ggplot2','reactable')
sapply(pkg, require, character = T)
# root.dir <- '/research/bsi/projects/staff_analysis/m216453/2022CyTOF/'
root.dir <- '/Users/m216453/Library/Mobile Documents/com~apple~CloudDocs/Documents/Mayo_project/2021_07_23_TinaCyTOF/'
list_df <- function(dea.res, list.name = "diff.marker.raw.p"){
  if(class(dea.res[[list.name]])=="list"){
    tb <- NULL
    for(n in 1:length(dea.res[[list.name]])){
      tb <- rbind(tb,dea.res[[list.name]][[n]])
    }
  }
  if(class(dea.res[[list.name]])=="array"){
    tb <- reshape2::melt(dea.res[[list.name]])
    colnames(tb) <- c('TimePoints','variables','celltypes','markers','value')
  }
  return(tb)
}
```


```{r, message=FALSE, message=F, warning=F, echo=FALSE, results='hide'}
##Subset the array data for T cell only or  BTNK cells separately
load('/Users/m216453/Library/Mobile Documents/com~apple~CloudDocs/Documents/Mayo_project/2021_07_23_TinaCyTOF/Data/2022_09_13/Anno.RData')
BTNK <- anno %>% dplyr::filter(compartment %in% c("T Cell", "NK Cell","B Cell"))
Tcell <- anno %>% dplyr::filter(compartment %in% c("T Cell"))
```

* Methods for Figure 5AB
    
    - Flow cytometry files were uploaded to the Astrolabe Cytometry Platform (Astrolabe Diagnostics, Inc) where transformation, cleaning (doublets, debris), labeling, and unsupervised clustering were done. Non-transformed data were downloaded and used in the statistical analysis. For differential abundance analysis (DAA), we applied total sum scaling normalization and asinhsqrt transformation to the count data. Then linear model was applied to compare the cell type abundance between bad vs. good response at different time points. For differential expression analysis (DEA), mean marker intensity was summarized per sample within each cell type. Then we calculated the log2 fold change of the mean marker intensity between C1, C3, C4 and baseline, respectively. Linear model was also applied to compare the marker intensity between bad vs. good response at different time points. For both DAA and DEA, a P value of <= 0.05 was considered significant and shown in the plots. Statistical analyses were performed using R software version 4.1.1.


# Figure 5A
```{r echo=F, error=TRUE, fig.height=14, fig.retina=4, fig.width=22, message=FALSE, warning=FALSE, include=T, results="asis"}
root.dir = '/Users/m216453/Library/Mobile Documents/com~apple~CloudDocs/Documents/Mayo_project/2021_07_23_TinaCyTOF/'

for(level in c('Assignment')){
  cat('----',level,'--- \n')
  for(type in c('BTNK')){
    cat('----',type,'--- \n')
    load(file.path(root.dir,'Data/2022_09_29', paste0(level, '.RData')))
    meta.dat0 <- meta.dat; cell.count0 <- cell.count
    load(paste0(root.dir,'Data/2022_09_29/', paste0(level, '_',type,'.RData'))) 
    meta.dat$SubjectID <- gsub('WWV54','WVW54',meta.dat$SubjectID)
    ## AG498 has been excluded in this RData
    meta.dat <- meta.dat0; cell.count <- cell.count0[,colnames(cell.count), drop =F]
    # meta2022 <- read.csv("~/Library/Mobile Documents/com~apple~CloudDocs/Documents/Mayo_project/2021_07_23_TinaCyTOF/Data/meta.csv") %>% 
    #   dplyr::filter(!is.na(TimePoint))
    # meta2022$PR2.n <- NA
    # meta2022[meta2022$BRAF == 'BRAFwt' & meta2022$PR4 %in% c('pCR','nc pCR','pPR'),'PR2.n'] <- 'good'
    # meta2022[meta2022$BRAF == 'BRAFwt' & meta2022$PR4 %in% c('pMR'),'PR2.n'] <- 'bad'
    # meta2022[meta2022$BRAF == 'BRAFm' & meta2022$PR4 %in% c('pCR','nc pCR'),'PR2.n'] <- 'good'
    # meta2022[meta2022$BRAF == 'BRAFm' & meta2022$PR4 %in% c('pMR','pPR'),'PR2.n'] <- 'bad'
    # meta.dat1 <- inner_join(meta.dat, meta2022 %>% dplyr::select(name, PR2.n))
    # tm <- load(file.path(root.dir,'Data/2022_09_29', paste0(level, '.RData')))
    # meta.dat <- meta.dat1
    # save(meta.dat, cell.count, cell.freq, cell.marker.mean, cell.marker.cvs, cell.marker.median, file = file.path(root.dir,'Data/2022_09_29', paste0(level, '.RData')))
    cell.freq <- cell.count/rowSums(cell.count)
    # load(paste0(root.dir,'Data/2022_09_29/', paste0(level, '_',type,'.RData')))
    # load(file.path(root.dir,'Data/2022_09_29', paste0(level, '.RData')))
    
    ## filter criterion
    if(level =='Assignment'){
      minp = 0.0;pct = 0
    }else{
      minp = 0.02;pct = 0.4
    }
    
    if(level == "Profiling"){
      ## collapse low abundance celltypes into 'other' group
      ind <- apply(cell.freq, 2, function(x) sum(x>minp) > (nrow(cell.freq) * pct))
      cell.freq1 <- cell.freq[,ind]
      other <- 1- rowSums(cell.freq1)
      cell.freq1 <- merge(cell.freq1, cbind(other), by = 0) %>% column_to_rownames('Row.names')
    }else{
      cell.freq1 <- cell.freq
    }
    
    col.name = 'file_name' ## name for join 2 dfs
    
    if(subsetBRAF ==T){
      meta.sub <- meta.dat %>% dplyr::filter(BRAF == BRAF.sub) %>% 
        dplyr::select(c(col.name,'TimePoint','SubjectID',grp.name)) %>% droplevels() 
    }else{
      meta.sub <- meta.dat %>% 
        dplyr::select(c(col.name,'TimePoint','SubjectID',grp.name)) %>% droplevels() 
    }
    
    plt <- as.data.frame(cbind(cell.freq1[meta.sub[,col.name],,drop =F])) %>% rownames_to_column(col.name) %>% inner_join(meta.sub, by = col.name)
    plt.m <- reshape2::melt(plt)
    
    plt.m$TimePoint <- gsub('BASELINE','base',plt.m$TimePoint)
    plt.m$TimePoint <- factor(plt.m$TimePoint, levels = c('base','C1','C3','C4'))
    plt.m$SubjectID <- gsub('WWV54','WVW54',plt.m$SubjectID)
    
    
    p1 <- ggplot(plt.m %>% dplyr::filter(!!as.name(grp.name)==unique(plt.m[,grp.name])[1]), aes(x = TimePoint, y = value, fill = variable)) +
      geom_bar(stat="identity") +
      facet_grid(. ~ SubjectID, scales = 'free',space = "free") +
      labs(y = paste0(unique(plt.m[,grp.name])[1],'(Proportion)'), x = '') +
      scale_fill_manual(values = c(brewer.pal(9,'Set3'),brewer.pal(9,'Set1'),brewer.pal(8,'Set2'),brewer.pal(8,'Dark2'))) +
      theme_bw() +
      theme(axis.text = element_text(color = 'black', size = 20),
            axis.title = element_text(color = 'black', size = 20),
            strip.text.x = element_text(color = 'black',size = 20),
            legend.text = element_text(color = 'black',size = 20),
            axis.text.x = element_text(color = 'black', angle = 90, size = 20, hjust = 1, vjust = 0.25),
            legend.position = 'bottom',
            legend.title = element_blank())
    
    p2 <- ggplot(plt.m %>% dplyr::filter(!!as.name(grp.name)==unique(plt.m[,grp.name])[2]), aes(x = TimePoint, y = value, fill = variable)) +
      geom_bar(stat="identity") +
      facet_grid(.~SubjectID, scales = 'free',space = "free") +
      # facet_wrap(PR2 ~ SubjectID, nrow = 2,  drop=TRUE) +
      labs(y = paste0(unique(plt.m[,grp.name])[2],'(Proportion)'), x = '') +
      scale_fill_manual(values = c(brewer.pal(9,'Set3'),brewer.pal(9,'Set1'),brewer.pal(8,'Set2'),brewer.pal(8,'Dark2'))) +
      theme_bw() +
      theme(axis.text = element_text(color = 'black', size = 20),
            axis.title = element_text(color = 'black', size = 20),
            strip.text.x = element_text(color = 'black',size = 20),
            legend.text = element_text(color = 'black',size = 20),
            axis.text.x = element_text(color = 'black', angle = 90, size = 20, hjust = 1, vjust = 0.25),
            legend.position = 'bottom',
            legend.title = element_blank())
    pp <- ggarrange(p1, p2, nrow = 2, common.legend = T)
    print(pp)
    # #ggsave(file = paste0("/Users/m216453/lu/Documents/Mayo_project/2021_07_23_TinaCyTOF/Code/Report_Summary/Folder2/Figure1A_",BRAF.sub,".pdf"), width =20, height = 12)
  }
}
```


```{r, message=FALSE, message=F, warning=F, echo=FALSE, results='hide'}
source(paste0(root.dir,'Code/Stats.R'))
source(paste0(root.dir,'Code/Func.R'))

cell.abund.obj <- cell.prop.obj <- cell.diff.obj <- list()
i = 0
for(level in c('Compartment','Assignment','Profiling')){
  i = i + 1
  output <- paste0(root.dir,'Result/',level)
  load(file.path(root.dir,'Data/2022_09_29', paste0(level, '.RData')))
  meta.dat$SubjectID <- gsub('WWV54','WVW54',meta.dat$SubjectID)
  
  if(subsetBRAF ==T){
    meta.dat <- meta.dat[meta.dat$file_name !="20201028_AG489_C1" & meta.dat$BRAF == BRAF.sub,]
  }else{
    meta.dat <- meta.dat[meta.dat$file_name !="20201028_AG489_C1", ]
  }
  cell.count <- cell.count[meta.dat$file_name,,drop =F]
  cell.freq <- cell.freq[meta.dat$file_name,,drop =F]
  # cell.count <- cell.count[,unique(BTNK[,tolower(level)])]
  ## cell.diff.obj 
  cell.count0 <- cell.count + 0.5
  size.factor <- rowSums(cell.count0)
  cell.freq <- cell.count0 / size.factor
  cell.freq.base <- cell.freq[meta.dat[meta.dat$TimePoint == 'BASELINE','file_name'], ]
  rownames(cell.freq.base) <- gsub("^([^_]*_[^_]*)_.*$", "\\1", rownames(cell.freq.base))
  base.name <- rownames(cell.freq.base)
  cell.freq.diff <- names <- NULL
  for(k in c('BASELINE','C1','C3','C4')){
    cell.time <- cell.freq[meta.dat[meta.dat$TimePoint == k,'file_name'], ]
    names <- c(names, rownames(cell.time))
    rownames(cell.time) <- gsub("^([^_]*_[^_]*)_.*$", "\\1", rownames(cell.time))
    cell.freq.diff <- rbind(cell.freq.diff, cell.time/cell.freq.base[rownames(cell.time),])
  }
  rownames(cell.freq.diff) <- names
  cell.freq.diff <- log2(cell.freq.diff) #log fold change from the baseline
  # cell.freq.diff.sqrt <- sign(cell.freq.diff) * sqrt(abs(cell.freq.diff ))
  
  cell.count.filter <- cell.count[,colSums(cell.count) > 0]
  cell.diff.obj[[paste0('level', i)]] <- cell.freq.diff[,colnames(cell.count.filter)]
  # cell.abund.obj[[paste0('level', i)]] <- cell.count
  cell.abund.obj[[paste0('level', i)]] <- cell.count.filter
  cell.prop.obj[[paste0('level', i)]] <- cell.freq[,colnames(cell.count)] # not filtered , all cells
}
names(cell.abund.obj) <- names(cell.prop.obj) <- names(cell.diff.obj)  <- c('Compartment','Assignment','Profiling')

meta.samples <- meta.dat[,c('file_name','TimePoint','BRAF','PR2','RunDate','patient_id','PR2.n','PR4.n')] %>% dplyr::rename(SubjectID = patient_id)
colnames(meta.samples)[1] <- 'sam.ids'
meta.samples$PR2.n <- as.factor(meta.samples$PR2.n)
meta.samples$PR4.n <- as.factor(meta.samples$PR4.n)
meta.samples <- within(meta.samples, BRAF <- relevel(BRAF, ref = 1))
meta.samples <- within(meta.samples, PR2 <- relevel(PR2, ref = 1))
```


```{r, message=FALSE, message=F, warning=F, echo=FALSE, results='hide'}
## perform differential abundance analysis, save daa.res of different timepoints into one file
run <- function(cell.obj, meta.samples, cell.plot.obj, method, 
                dir.l1, dir.l2 = 'observed',
                cell.levels = c('Compartment','Assignment','Profiling'),
                grp.name, adj.name, timepoints =c('BASELINE','C1','C3','C4') ,BRAF.sub,transform,
                subsetBRAF = F,filter = T, feature.dat.type = 'count',diff = '',labels.y = 'Proportion'){
  for(cell.level in cell.levels){ #
    output <- paste0(dir.l1, dir.l2, '/',cell.level,'/')
    if(!dir.exists(output)){dir.create(output)}
    
    output1 <- paste0(output,'testing',grp.name,'_adj',paste0(adj.name,collapse = '.'),'_subset',subsetBRAF,diff,'/')
    if(!dir.exists(output1)){dir.create(output1)}

    daa <- NULL
    for(time in timepoints){
      if(subsetBRAF == T){
        meta <- meta.samples %>% dplyr::filter(meta.samples$TimePoint == time & meta.samples$BRAF==BRAF.sub)
      }else{
        meta <- meta.samples %>% dplyr::filter(meta.samples$TimePoint == time)
      }
      
      data.obj <- list()
      for(k in names(cell.obj)){
        cell.sub <- cell.obj[[k]]
        data.obj[[k]] <- cell.sub[meta$sam.ids,]
      }
      
      set.seed(123)
      daa.res <- perform_DAA(data.obj = data.obj, meta.samples = meta, cell.level = cell.level,
                             grp.name, adj.name, strata = NULL, method = method, filter= filter,
                             transform = transform,
                             feature.dat.type, cell.depth = 0, output = NULL)
      daa <- rbind(daa, daa.res$res %>% rownames_to_column('cells') %>% dplyr::mutate(time = time))

    }
    save(daa, file = paste0(output1,'daa.res.',method,'.RData')) ## all timepoints daa.res will be saved into one file, for later plotting purpose
    
    # load(paste0(output1,'daa.res.',method, '.RData'))
    if(method == 'LinDA'){
      daa.table <- daa %>% dplyr::select(cells,log2FoldChange, lfcSE,pvalue, padj, time) %>% 
      dplyr::mutate(log2FoldChange = round(log2FoldChange,2), lfcSE = round(lfcSE,2),
             pvalue = round(pvalue,2), padj = round(padj,2)) %>% arrange(pvalue)
    }
    
     if(method == 'lm'){
      daa.table <- daa %>% dplyr::mutate(coef = round(coef,2), sd.err = round(sd.err,2),
             pvalue = pvalue, padj = padj) %>% arrange(pvalue)
    }
    
    col <- c(sig = 'red', ns = 'grey')
    vol.df <- daa %>% dplyr::mutate(label = ifelse(pvalue <= 0.05, 'sig','ns'), log10pval = -log10(pvalue))
    ct <- sum(daa$pvalue<= 0.05)
    pct <- round(ct/nrow(daa) * 100)
    if(method == 'lm'){x.col = 'coef'}
    if(method == 'LinDA'){x.col = 'log2FoldChange'}
    v <- ggplot(vol.df, aes(x = !!as.name(x.col), y = log10pval, color = label)) +
      geom_point(size = 1) +
      scale_color_manual(values = col) +
      labs(y = '-log10P') +
      theme_bw() +
      theme(legend.position = "none",
            axis.text = element_text(size = 16),
            axis.title = element_text(size = 16)) +
      annotate("text", x=0, y=max(vol.df$log10pval) * 0.95, 
               label= paste0(ct, '(',pct, '% tested) markers raw.p <= 0.05 in all timepoints')) 
    daa <- daa %>% dplyr::mutate(pvalue = round(pvalue,2), padj = round(padj,2)) 
    timeplot.rawp <- plot_timeseries_boxplot(cell.plot.obj, meta.samples, daa, 
                                             cell.level=cell.level, grp.name, labels.y = labels.y,
                                             add.smooth =T,
                                             plot.type = 'pointplot', cutoff = 0.2, cutoff.pval = 0.05, 
                                            show.p = 'p', root.dir= NULL)
    timeplot.box <- plot_timeseries_boxplot(cell.plot.obj, meta.samples, daa, 
                                             cell.level=cell.level, grp.name, labels.y = labels.y,
                                             add.smooth =T,
                                             plot.type = 'boxplot', cutoff = 0.2, cutoff.pval = 0.05, 
                                            show.p = 'p', root.dir= NULL)
    timeplot.line <- plot_timeseries_boxplot(cell.plot.obj, meta.samples, daa, 
                                             cell.level=cell.level, grp.name, labels.y = labels.y,
                                             add.smooth =T,
                                             plot.type = 'lineplot', cutoff = 0.2, cutoff.pval = 0.05, 
                                            show.p = 'p', root.dir= NULL)
  }
  return(list(tb = daa, volcano = v, timeplot.rawp = timeplot.rawp, timeplot.box = timeplot.box, timeplot.line = timeplot.line))
}

```

# Figure 5B
```{r, message=FALSE, message=F, warning=F, echo=FALSE, results='hide'}
CD4 <- c("CD4+ T Cell (Central Memory)","CD4+ T Cell (Effector Memory)","CD4+ T Cell (EMRA)","CD4+ T Cell (Naive)")
CD8 <- c("CD8+ T Cell (Central Memory)","CD8+ T Cell (Effector Memory)","CD8+ T Cell (EMRA)","CD8+ T Cell (Naive)")
dir.l1 <- '/Users/M216453/Documents/Mayo_project/2021_07_23_TinaCyTOF/Result/2022_12_14_FinalSummary/DAA/'
### Figure 1B
cell.prop.obj$CD48 <- cbind(`CD4+ T cells`= rowSums(cell.prop.obj$Assignment[,CD4]), 
                            `CD8+ T cells`=rowSums(cell.prop.obj$Assignment[,CD8]),
                            `B cell`= (cell.prop.obj$Compartment[,'B Cell']),
                            `NK cell`= (cell.prop.obj$Compartment[,'NK Cell']),
                            `gd T Cell` =(cell.prop.obj$Assignment[,'gd T Cell']))
ree <- NULL
ree <- run(cell.obj = cell.prop.obj, cell.plot.obj = cell.prop.obj, 
           dir.l1 = dir.l1, dir.l2 = 'mean_observed', 
           cell.levels = 'CD48',timepoints =c('BASELINE','C1','C3','C4'),
           transform = 'asinhsqrt',
           feature.dat.type = 'other', filter = F, method = 'lm', subsetBRAF = subsetBRAF, BRAF.sub = BRAF.sub,
           meta.samples = meta.samples, grp.name = grp.name, adj.name = adj.name)
res <- ree$tb[,c('cells','pvalue','time')] %>% spread(time, pvalue) %>% column_to_rownames('cells')
reactable(res)
res0 <- t(res)
```

```{r echo=F, error=TRUE, fig.height=6, fig.retina=4, fig.width=14, message=FALSE, warning=FALSE, include=T, results="asis"}
plt <- list()
for(celltype in colnames(cell.prop.obj$CD48)){
  data <- as.data.frame(cell.prop.obj$CD48[,c(celltype), drop =F]) %>% rownames_to_column('sam.ids')
  meta.tmp <- meta.samples[,c('sam.ids','TimePoint','PR4.n')]
  colnames(meta.tmp)[3] <- 'Response'
  data1 <- full_join(data, meta.tmp) 
  p <- as.data.frame(res0[,celltype,drop =F])%>% rownames_to_column('TimePoint')
  colnames(p)[2] <- 'p'
  data2 <- data1 %>% left_join(p)
  data2$TimePoint <- paste0(data2$TimePoint, '(p=',data2$p,')')
  data2$Response <- gsub('good',paste0(as.character(expression("\u2264")),10),data2$Response)
  data2$Response <- gsub('bad','>10',data2$Response)
  data2 <- within(data2, Response <- factor(Response, levels = c(paste0(as.character(expression("\u2264")),10),'>10')))
  plt[[celltype]] <- ggplot(data2, aes(y = (!!as.name(celltype))*100, x = Response, color =!!as.name('Response'))) +
    geom_boxplot() +
    geom_jitter(position = position_jitterdodge(0.5)) +
    facet_wrap(~TimePoint, nrow =1) +
    scale_color_manual(values = c("≤10"=brewer.pal(8,'Set1')[2],">10"=brewer.pal(8,'Set1')[1])) +
    theme_bw() +
    theme(axis.title = element_text(size = 14),
          panel.grid = element_blank(),
          strip.text.x = element_text(size = 12),
          axis.text = element_text(size = 12, color = 'black'),
          legend.position = 'none',
          plot.title = element_text(hjust = 0.5)) +
    labs(x = 'Pathologic response (% viable tumor cells)', y = '% of total cells') +
    ggtitle(celltype)
}
pp <- ggarrange(plt$`CD4+ T cells`, plt$`CD8+ T cells`, plt$`B cell`, plt$`NK cell`, plt$`gd T Cell`)
```


```{r echo=F, error=TRUE, fig.height=6, fig.retina=4, fig.width=10, message=FALSE, warning=FALSE, include=T, results="asis"}
dir.l1 <- '/Users/M216453/Documents/Mayo_project/2021_07_23_TinaCyTOF/Result/2022_12_14_FinalSummary/DAA/'
ree <- NULL
ree <- run(cell.obj = cell.abund.obj, cell.plot.obj = cell.prop.obj, dir.l1 = dir.l1, dir.l2 = 'mean_observed', 
           cell.levels = 'Assignment',timepoints =c('BASELINE','C1','C3','C4'),
           transform = 'TSS.asinhsqrt',
           feature.dat.type = 'other', filter = F, method = 'lm', subsetBRAF = subsetBRAF, BRAF.sub = BRAF.sub,
           meta.samples = meta.samples, grp.name = grp.name, adj.name = adj.name)
res <- ree$tb[,c('cells','pvalue','time')] %>% spread(time, pvalue) %>% dplyr::filter(cells %in% c(CD4,CD8))%>% column_to_rownames('cells')
reactable(res)
res0 <- t(res)
```

```{r echo=F, error=TRUE, fig.height=6, fig.retina=4, fig.width=10, message=FALSE, warning=FALSE, include=T, results="asis"}
plt2 <- list()
for(celltype in c(CD4, CD8)){
  data <- as.data.frame(cell.prop.obj$Assignment[,c(celltype), drop =F]) %>% rownames_to_column('sam.ids')
  meta.tmp <- meta.samples[,c('sam.ids','TimePoint','PR4.n')]
  colnames(meta.tmp)[3] <- 'Response'
  data1 <- full_join(data, meta.tmp) 
  p <- as.data.frame(res0[,celltype,drop =F])%>% rownames_to_column('TimePoint')
  colnames(p)[2] <- 'p'
  data2 <- data1 %>% left_join(p)
  data2$TimePoint <- paste0(data2$TimePoint, '(p=',data2$p,')')
    data2$Response <- gsub('good',paste0(as.character(expression("\u2264")),10),data2$Response)
  data2$Response <- gsub('bad','>10',data2$Response)
  data2 <- within(data2, Response <- factor(Response, levels = c(paste0(as.character(expression("\u2264")),10),'>10')))

  plt2[[celltype]] <- ggplot(data2, aes(y = (!!as.name(celltype))*100, x = Response, color =!!as.name('Response'))) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(position = position_jitterdodge(0.5)) +
    facet_wrap(~TimePoint, nrow =1) +
    scale_color_manual(values = c("≤10"=brewer.pal(8,'Set1')[2],">10"=brewer.pal(8,'Set1')[1])) +
    theme_bw() +
    theme(axis.title = element_text(size = 14),
          panel.grid = element_blank(),
          strip.text.x = element_text(size = 12),
          axis.text = element_text(size = 12, color = 'black'),
          legend.position = 'none',
          plot.title = element_text(hjust = 0.5)) +
    labs(x = 'Pathologic response (% viable tumor cells)', y = '% of total cells') +
    ggtitle(celltype)
}

ggarrange(plt$`CD4+ T cells`, plt$`CD8+ T cells`, plt$`B cell`, plt$`NK cell`, plt$`gd T Cell`,plt2$`CD8+ T Cell (Central Memory)`)
```




# Figure 5C
```{r}
level = 'Compartment'
tmp <- load(paste0("/Users/m216453/Library/Mobile Documents/com~apple~CloudDocs/Documents/Mayo_project/2021_07_23_TinaCyTOF/Data/2022/agg_trans/",level,".RData"))
cell.marker.mean[is.na(cell.marker.mean)] = 0
meta.dat$SubjectID <- gsub('WWV54','WVW54',meta.dat$SubjectID)

if(subsetBRAF == T){
  meta.samples <- meta.samples[meta.samples$sam.ids != "20201028_AG489_C1" & meta.samples$BRAF == BRAF.sub,,drop =F]
}else{
  meta.samples <- meta.samples[meta.samples$sam.ids != "20201028_AG489_C1" ,,drop =F]
}

files <- meta.samples$sam.ids
celltypes.T <- 'T Cell'

cell.marker.mean1 <- 
  array(NA, c(length(files), 1, length(dimnames(cell.marker.mean)[[3]])), 
        dimnames = list(files, c('T Cell'),dimnames(cell.marker.mean)[[3]]))
for(file in files){
  tmp1 <- as.data.frame(cell.marker.mean[file,,][c(celltypes.T),,drop =F]) %>% 
    dplyr::mutate(grp = gsub('T Cell.*','',rownames(.))) %>% 
    group_by(grp) %>% 
    summarise_if(is.numeric, sum, na.rm = TRUE) %>% column_to_rownames('grp') %>% as.matrix
}

cell.marker.mean.agg.Tcell <- cell.marker.mean1
cell.marker.mean.agg.fc.Tcell  <- lfc.func(cell.marker = cell.marker.mean.agg.Tcell)$cell.marker.fc

```

```{r}
celltypes <- c("T Cell")
dea.res1 <- NULL
dea.res1 <- perform_DEA_CyTOF(data = cell.marker.mean.agg.Tcell, celltypes = celltypes, 
                              sam.col = 'sam.ids', method = 'lm',transform = 'log2',
                              BRAF.sub =BRAF.sub, subsetBRAF = subsetBRAF,
                              meta.dat = meta.samples, timepoints = c('BASELINE','C1','C3','C4'), 
                              variables = grp.name,adj.name = adj.name)
df <- dea2tb(dea.res1)$df
pp1 <- plot_time(cell.marker.data = (cell.marker.mean.agg.fc.Tcell),
                dea.res = df %>% dplyr::mutate(raw.p = round(raw.p,2), fdr = round(fdr,2)),
                add.smooth =T,timepoints = c('BASELINE','C1','C3','C4'),
                meta.dat = meta.samples, p.type = 'raw.p', cutoff = 0.05,
                plot.type = 'lineplot', sam.col = 'sam.ids',
                celltypes = celltypes, grp.name = grp.name)
print(pp1$`T Cell_PD-L1`)
```
